---
title             : "An Antidote to Corruption: When Social Environments and Partner Choices Do More Harm Than Good"
shorttitle        : "Choosing Corrupt Collaboration"

author: 
  - name          : "Sebastian Simon"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "s.simon.2@umail.leidenuniv.nl"

affiliation:
  - id            : "1"
    institution   : "Social, Economic, and Organizational Psychology, Leiden University"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: Goes here.
  
keywords          : "Rule following, partner choice, social environment, norms"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
keep_tex          : yes
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

Cooperation such as rule-following, for instance, is key to the functioning of societies. People can be prosocial, follow rules, and therewith further the interests of a group as a whole. Cooperation, however, is not ALWAYS prosocial per se. Instead, people can defect as a collective when they face higher benefits for doing so, a concept called corrupt collaboration. This form of corruption only benefits those defecting and comes at the expense of others. While the consequences of prosocial behavior are largely considered positive, the consequences of corrupt collaboration can be disastrous. [ex]. We, therefore, have to investigate when groups cooperate for prosocial reasons and when groups corrupt. 

# From Costly Signaling to Reputation Management

When individuals cooperate, all involved parties have to consider both the costs and the benefits. Importantly, costs can be exploitation attempts, risks that are better avoided if possible. To shield themselves from being exploited, individuals evaluate how trustworthy others are. [ex]. Individuals readily signal their trustworthiness. For instance, charity [ex]. Importantly, signals are considered reliable when they are rather costly for some to perform or others do not possess the qualities in the first place. [ex]. Therefore, cooperation requires trustworthiness among the parties involved as a safety mechanism. 

Reciprocity is a well-documented way of safeguarding cooperation and is most likely when group members know each other. Evolutionary mechanisms, for instance, point to kin selection where individuals are more likely to reciprocate help from relatives than non-relatives, and group selection where individuals serve the existence of a group and, thereby, reciprocate. However, in the face of globalization, individuals interact less frequently and do not necessarily know each other which, in turn, gradually removes incentives to reciprocate. [ex]. Therefore, cooperation in anonymous groups requires another mechanism. 

Indirect reciprocity is such a way of safeguarding cooperation when group members do not know each other and social interactions are limited to one. This form of reciprocity occurs when an observer helps an individual because the observer saw that individual helping someone else. [ex]. Indirect reciprocity is the underlying mechanism of reputation, the public opinion of a person [def]. [ex]. Therefore, individuals use the reputation of others to evaluate whether they are a good interaction partner. 
  
Reputation has been described as a universal currency for social interactions, and individuals treat it as such. The concept of the updating strategy is based on individuals keeping score of others' and their signals. Within this concept, various models conceptualize reputation differently. For example, the standing strategy explains how, in the context of helping, all group members already start with a good reputation score and can loose their good standing if they fail to help, thus, reducing their score. However, this model did not prevail over the scoring strategy. The scoring model holds that group members' reputation scores rise and fall with their acts of help. [ex]. Such updating strategies are reasons for individuals to deliberately manage their reputation, also called reputation management. [ex]. Therefore, individuals manage and care about others' impressions of them. 

# Partner Choice in Biological Markets

Reputation management is vital for biological markets where individuals compete to choose and be chosen as interaction partners. Unlike economic markets where goods and services are traded, biological markets focus on the ability and willingness to benefit others. Hunter-gatherer tribes, peers [ex]. Importantly, the willingness to benefit is more important than the ability to do so. For instance, [blabla] found that generosity prevails over productivity [ex]. Reputation, therefore, plays a vital role in biological markets.

However, in a market, supply commonly follows demand. Specifically, individuals compete to be chosen but also choose who they want to interact with such as their partners, friends, and colleagues at work, a concept we refer to here as partner choice. Importantly, reputation helps individuals choosing good interaction partners [ex]. Therefore, if competitors do not have a desirable reputation they will not be chosen as partners and will thus miss out on potential benefits [ex]. Partner choice, therefore, creates selection pressures for competitors.  

Which qualities are considered desirable is conditional on the goals of the individual choosing. For instance, in a prosocial environment, signals of prosociality are preferred [ref] and, therefore, lead to cooperation. [ex]. However, individuals also cooperate to corrupt. In fact, in a corrupt environment, people have been shown to prefer corrupt collaboration, the joint violation of rules for the sake of gaining benefits. [ex]. H1.  

# Cooperation Within and Across Environments

Both prosociality and corrupt collaboration can increase in frequency in the form of social norms over time. Social norms are commonly defined as common behaviors that either should be acted out (i.e., injunctive norms) or are indeed acted out (i.e., descriptive norms). [ex]. For instance, previous research has examined social norms in prosocial environments with prosocial behavior increasing future acts of prosociality over time. For instance, [blabla] competitive altruism. Similar to prosocial behavior, corruption has been shown to breed corruption. [blabla] has shown that. H2.

Moreover, social environments have been shown to interact when individuals switch from one environment to another. Specifically, when participants were exposed to a prosocial environment followed by a dishonest environment, the participants were more prosocial in the corrupt environment than if they had not been exposed to the prosocial environment in the first place. For instance, [blabla] has shown that cooperation stayed high when reputation. Also, when the environments were reversed, participants became more corrupt. [blablabla] has shown that. H3.

# Methods

# Results

# Discussion 

- As deciders repeatedly completed the die-rolling task in previous experimental blocks, we would expect their responses to differ from selectors’ responses. 
- feedback from participants
  - terms 'selector' and 'decider' confusing 
- methods
  - die rolling game in the end confounding the svo
- discrepancy between injunctive (what we told participants to do) and descriptive norm (what they actually did)
- theoretical implications
  - typically more cooperation 
- practical implications
  - from experimental setup
  - from 

# Conclusion

\newpage
# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup

\newpage
# Appendix A

Syntax goes here. 

\newpage
# Appendix B

## Information Brochure

Dear participant, this brochure provides you with information about the type and methods of the study in which you are about to participate. It is therefore important that you read this document closely. 

### Purpose of the Study

People constantly make decisions, sometimes to improve their situation and sometimes to prevent it from worsening. In this study we will let you make a series of decisions, in which you can increase or decrease your starting capital. Whatever you have earned from your decisions during the task will be paid out to you in the end. We expect that the decision task is involving and that we get a good insight into the kind of investments you make.
 
### What is going to happen?

After you have read this introduction and signed the informed consent form, you will be briefed and trained in the task. It is important for you to know that you can leave the experiment at any point without providing a justification and without consequences. In this experiment, you will make a number of decisions. Each time, performance and earnings will be measured. At the end of this study, you will receive debriefing and eventual earnings. We will not provide your personal information to anybody else, only use these for scientific purposes, and will only report results averaged over all participants and not about individual cases.

### Financial reward

In this experiment, you participate in 1 session of about 60 minutes. You will receive a participation fee of 6,50 Euros (or 2 credits if you prefer) independent of your performance. In addition, depending on your decisions you may earn up to 6,50 Euros for your participation. You may thus earn up to 13 Euros in total. Your earnings will be calculated after the conclusion of the experiment and paid out to you after the second session.
 
### Voluntary participation

If you now decide not to participate in this experiment, this shall have no consequences for you. If you decide during the experiment to withdraw from the study, this shall have no consequences for you. In addition, up to 24 hours after the study you can still withdraw your consent for use of your personal information. You can thus withdraw your participation at any point. You are free to do so without providing any justification. If you now or within the next 24 hours want to withdraw your consent, your personal information will be removed from our database.
 
### Confidentiality of study results

All information from this study will remain coded. The principal investigator has no insight into your identity and will transfer any sum to be transferred to you to the research assistants in sealed envelopes. Thus, the experimenters do not know how much money you earned.

### Debriefing

At the end of this session, you will receive a short summary of the purpose of this study. You can always direct questions about the experiment to the experimenters or per email to Dr. Jörg Gross (j.a.j.gross@fsw.leidenuniv.nl).

### Informed Consent 

This study involves the reading of instructions and making a series of decisions that can affect your payment. All instructions, decisions, and questionnaires will be presented to you on the computer. At the end of the experiment, you will receive a debriefing with background information on the study, along with the additional earnings you obtained during the experiment.  The additional earnings depend on your decisions and can range between 2 and 6,50 Euros. How much you have earned will be paid out to you in cash after the session.

The study involves one session and you will be compensated 6,50 Euros or 2 credits. In addition, you can earn more during the study itself. All measures taken in this study are for scientific purposes only and will be stored in a coded way. Participation is voluntary and at your own discretion. This means that you can withdraw from the study at any time and without having to explain or justify why. You will still receive the show-up fee of 6,50 Euros or 2 credits. All information collected during this study is confidential and the data will be stored in such a way that responses cannot be traced back to your identity. The study is coordinated by Dr. Jörg Gross (j.a.j.gross@fsw.leidenuniv.nl). Questions or complaints can be addressed to him.

I herewith confirm that I have read and understood the information brochure and that I consent with participating in this study.

\newpage
# Appendix C
## Instructions

Welcome to the experiment! Below you will find detailed information about the study and a short test to check whether you understood the general setup. It is therefore important that you read the instructions closely. Click the blue headings to collapse the subsections. There is no deception and no hidden information in this study. Please do not hesitate to call the experimenter if anything remains unclear to you. Note: Tick the check boxes in the subsections below to show that you have read and understood the instructions. Otherwise, you will not be able to proceed. 

In this study, you will be assigned to one of two roles and you will remain in this role throughout the experiment. You will either be playing in the role of the "selector" or in the role of a "decider". In total, there is one selector and there are three deciders. You will find out about your role at the start of the experiment. 

## Part 1

This study consists of two parts. Below, we will explain the first part in detail. After you have completed the first part of the experiment, we will give you instructions about the second part. At the end of the experiment, one round of part 1 will be selected randomly by the computer. Since you do not know which round will count for real, you should treat each round independently and as if every round is the one that counts. The points you earn in a round will be converted to money at a conversion rate of 100 points = 1 Euro. Hence, your decisions have real consequences for your earnings and, potentially, the earnings of other participants. You will start with 0 points and if your point total is below 650 points at the end of the experiment, you will still get paid 6.50 Euros. Therefore, you can earn a bonus if your point total is above 650 points. 

### Stage 1. 

The first part of the study consists of 15 rounds. Each round has three stages. Each decider will decide how to allocate 15 balls between two buckets on the computer screen. The deciders' task is to put each ball, one-by-one, into one of the two buckets: the blue bucket or the yellow bucket. For each ball the decider puts in the blue bucket he or she will receive 5 points and for each ball the decider puts in the yellow bucket he or she will receive 15 points. The rule is to put the balls in the blue bucket. The deciders' payments in this stage will be based on the sum of the points of the blue bucket and the yellow bucket. The selector will not take part in stage 1. 

### Stage 2. 

The selector will start by receiving 450 points. The selector will then learn about the decisions of all three deciders. Specifically, the selector will be told how many balls each decider placed in the blue bucket. The selector can then choose which decider to interact with for stage 3. The selector has to select at least one decider to interact with but can also choose to interact with two deciders in stage 3 - or even with all three. For every decider that the selector chooses, the selector has to pay a cost of 150 points. If a decider is not selected for stage 3, he or she will skip this stage, wait for the others to finish, and not earn be able to earn more. Importantly, the selector will not be able to identify the deciders across rounds, but only learn about their behavior in stage 1 (the bucket task). Specifically, the selector will be told how many balls each decider placed in the blue bucket. 

### Stage 3. 

If a decider is selected as interaction partner, he or she will receive 500 points. The decider is then asked how many points he or she wants to keep and how many points he or she wants to give to the selector. Hence, the decision of the decider determines the earnings of the decider as well as the earnings of the selector in this stage. After the decider has made his or her decision, the selector will learn about the outcome. 

### Feedback in part 1. 

After stage 3, the round is over, and you will receive a summary of this round. In the role of the decider, you receive a summary of: (a) your payoff from stage 1, (b) whether you were selected as interaction partner for stage 3 (c) how many points you decided to keep for yourself and give to the selector (d) your total sum of points you earned in this round In the role of the selector, you receive a summary of: (a) the deciders you chose as interaction partners for stage 3 (b) how many points the deciders you interacted with decided to keep for themselves and give to you (c) your total sum of points you earned in this round. Then, you move to the next round starting with stage 1.

## Part 2 

In this part, everything will stay the same as in part 1, except for stage 3 and the feedback. You will also stay in your role (decider or selector) from part 1. Again, this part consists of 15 rounds. For your convenience, we repeat the instructions for stage 1 and 2 below. Again, click the blue headings to collapse the subsections. Note: Tick the check boxes in the subsections below to show that you have read and understood the instructions. Otherwise, you will not be able to proceed. 

In this stage the selected deciders will use the die and the cup. The deciders have to roll the die using the cup, peek under the cup, and report the die-roll outcome. The payoff for the decider and the selector will be determined by the result that the decider reports. Specifically: If a decider reports a 1, both the decider and the selector will earn 0 points. If a decider reports a 2, both the decider and the selector will earn 50 points. If a decider reports a 3, both the decider and the selector will earn 100 points. If a decider reports a 4, both the decider and the selector will earn 150 points. If a decider reports a 5, both the decider and the selector will earn 200 points. If a decider reports a 6, both the decider and the selector will earn 250 points.

### Feedback in part 2. 

After stage 3, the round is over, and you will receive a summary of this round. In the role of the decider, you receive a summary of: (a) your payoff from stage 1 (b) whether you were selected as interaction partner for stage 3, and (c) your die-roll report and how many points you and the selector earned, accordingly In the role of the selector, you receive a summary of: (a) the deciders you chose as interaction partners for stage 3 and (b) the die-roll report and resulting earnings for each decider you interacted with. Then, you move on to the next round starting with stage 1.

## Debriefing

In this study, you were part of a four-person group and made a series of decisions that could affect your final payoff. In one part of the experiment you were confronted with a rule of how to make decisions. We were interested in how many people and to what extent they follow this rule. In another part, one person of your group had to decide who to interact with based on the decisions that you and others made before. We are interested in when people choose to interact with others based on others’ previous decisions in the rule-task. In the last part, there were two contexts: you were assigned to either report the rolls of a die or divide money among yourself and a partner. If you had to report the rolls of a die, by not reporting truthfully, you were able to earn more money. We are interested to what extent over-reporting in this task is related to following the rule in the first task and being chosen as a partner. If you had to divide money among yourself and a partner, by giving more to yourself, you were able to earn more money. We are interested to what extent people make fair allocations and how this is related to getting chosen as a partner and signaling to follow rules in the rule-task. The study will help us better understand when and why individuals choose to interact with other people and follow or break rules. 

The study did not involve any deception – everything that was told to you did in fact happen and/or will be implemented upon completion of the study. For further information, please contact the coordinator the study, Dr Jörg Gross (j.a.j.gross@fsw.leidenuniv.nl). Thank you for your participation!

